Sparse Function-space Representation \\ of Neural Networks

### 2023-05-17 paper/main.tex
updated rl
tyos

% Introduce the laplace GGN + shortcomings
\textbf{Function-space methods}
Function-space perspectives on uncertainty in neural networks often use a Laplace-GGN approximation \citep{daxberger2021laplace}, that linearize a trained neural network around MAP weights and approximates the neural network's Hessian using the generalized Gauss--Newton approximation (GGN, \cite{botev2017practical}). 
%\todo{Could modify the sentence a bit, maybe "The Laplace-GGN approximates the Hessian by using..."}
%The Hessian can be efficiently approximated using the generalized Gauss--Newton approximation . 
While efficient, it suffers from cubic scaling in parameter count, necessitating approximations like Kronecker factorisation \cite{martens2015optimizing, ritter2018kfac}. \todo{Reads a bit repetitive to start two sentences in a row by This, could maybe be combined in one.} This linear model (linear in the weights) refines predictions and provides uncertainty estimates \citep{immer2021scalable}. This model has a convenient interpretation as a GP \cite{immer2021scalable, khan2019approximate, maddox2021fast}.
%
However, the GP's cubic $\mathcal{O}(n^3)$ scaling with data points $n$ requires costly approximations, often resorting to using data subsets \cite{immer2021scalable}. In the GP community, sparse approximations have mitigated this scaling issue (\eg, \cite{titsias2009variational,hensman2013gaussian}), but combining neural network linearization with sparse methods remains \todo{Maybe unexplored instead of unclear?} unclear without resorting to subset methods or separately retraining a GP model \cite{ortega2023variational} \todo{Could add something on subset methods here, for example 'which do not retain information from the full dataset' or something like that}. Our work addresses this by utilizing the GP's dual parameters \cite{csato2002sparse}, previously applied to non-conjugate likelihood models \cite{adam2021dual}.

%wang2019exact
%
%However, the GP introduces a cubic $\mathcal{O}(N^3)$ scaling in the number of data points $N$. Previous approaches apply crude approximations to alleviate the computational costs, \eg, by considering only a subset of the training data \cite{immer2021scalable}. In the GP community, the scaling problem is not seen as an issue any more due to efficient and stochastic methods for sparse approximations (\eg, \cite{hensman2013gaussian,wang2019exact}). However, it is not entirely clear how to combine the linearization of the neural network with sparse methods \cite{titsias2009variational} without falling back to crude subset methods. We tackle this problem by framing the problem in the dual parameters \cite{csato2002sparse} of the GP that has previously been used for non-conjugate likelihood models in GPs \cite{adam2021dual}.

@@

\textbf{Uncertainty quantification in RL}
A key challenge in RL is balancing the trade-off between exploration and exploitation \cite{sutton2018reinforcement}.
%That is, should an agent select actions that it knows will lead to high reward %(exploitation), or should it
%select new ones in hope to discover actions leading to higher reward (exploration).
One promising direction to balancing this trade-off is to model the uncertainty associated with a learned transition dynamics model and use it to guide exploration.
Prior work has taken an expectation over the dynamic model's posterior \cite{deisenroth2011pilco,kamtheDataEfficient2018,chuaDeepReinforcementLearning2018},
sampled from it (akin to Thompson sampling but referred to as posterior sampling in RL)
\todo{cite thopmson sampling paper dearden et al 1999}
\cite{osbandMoreEfficientReinforcement2013},
and used it to implement optimism in the face of uncertainty using upper confidence bounds \cite{curiEfficient2020}.
\todo{cite UCRL paper}
No single strategy has emerged as a go-to method and we highlight that these strategies are only as good as their uncertainty estimates.
Previous approaches have used GPs \cite{deisenroth2011pilco,kamtheDataEfficient2018},
ensembles of neural networks \cite{curiEfficient2020,chuaDeepReinforcementLearning2018}
and variational inference \cite{galImproving2016,houthooftVIME2017}. \todo{Could specify here how the previous methods have used these models}
%However, there has been little leveraging

>>>>

% Introduce the laplace GGN + shortcomings
\textbf{Function-space methods}
Function-space perspectives on uncertainty in neural networks often use a Laplace-GGN approximation \citep{daxberger2021laplace}, which linearizes a trained neural network around MAP weights and approximates the neural network's Hessian using the generalized Gauss--Newton approximation (GGN, \cite{botev2017practical}).
%\todo{Could modify the sentence a bit, maybe "The Laplace-GGN approximates the Hessian by using..."}
%The Hessian can be efficiently approximated using the generalized Gauss--Newton approximation . 
While efficient, it suffers from cubic scaling in parameter count, necessitating approximations like Kronecker factorisation \cite{martens2015optimizing, ritter2018kfac}. This linear model (linear in the weights) refines predictions and provides uncertainty estimates \citep{immer2021scalable}. It has a convenient interpretation as a GP \cite{immer2021scalable, khan2019approximate, maddox2021fast}.
%
However, the GP's cubic $\mathcal{O}(N^3)$ scaling with data points $N$ requires costly approximations, often resorting to using data subsets \cite{immer2021scalable}. In the GP community, sparse approximations have mitigated this scaling issue (\eg, \cite{titsias2009variational,hensman2013gaussian}), but combining neural network linearization with sparse methods remains unclear without resorting to subset methods or separately retraining a GP model \cite{ortega2023variational}. Our work addresses this by utilizing the GP's dual parameters \cite{csato2002sparse}, previously applied to non-conjugate likelihood models \cite{adam2021dual}.

% \todo{Could add something on subset methods here, for example 'which do not retain information from the full dataset' or something like that}.
%wang2019exact
%
%However, the GP introduces a cubic $\mathcal{O}(N^3)$ scaling in the number of data points $N$. Previous approaches apply crude approximations to alleviate the computational costs, \eg, by considering only a subset of the training data \cite{immer2021scalable}. In the GP community, the scaling problem is not seen as an issue any more due to efficient and stochastic methods for sparse approximations (\eg, \cite{hensman2013gaussian,wang2019exact}). However, it is not entirely clear how to combine the linearization of the neural network with sparse methods \cite{titsias2009variational} without falling back to crude subset methods. We tackle this problem by framing the problem in the dual parameters \cite{csato2002sparse} of the GP that has previously been used for non-conjugate likelihood models in GPs \cite{adam2021dual}.

@@

\textbf{Uncertainty quantification in RL}
A key challenge in RL is balancing the trade-off between exploration and exploitation \cite{sutton2018reinforcement}.
%That is, should an agent select actions that it knows will lead to high reward %(exploitation), or should it
%select new ones in hope to discover actions leading to higher reward (exploration).
One promising direction to balancing this trade-off is to model the uncertainty associated with a learned transition dynamics model and use it to guide exploration.
Prior work has taken an expectation over the dynamic model's posterior \cite{deisenroth2011pilco,kamtheDataEfficient2018,chuaDeepReinforcementLearning2018},
sampled from it (akin to Thompson sampling but referred to as posterior sampling in RL)
\todo{cite thopmson sampling paper dearden et al 1999}
\cite{osbandMoreEfficientReinforcement2013},
and used it to implement optimism in the face of uncertainty using upper confidence bounds \cite{curiEfficient2020}.
\todo{cite UCRL paper}
No single strategy has emerged as a go-to method and we highlight that these strategies are only as good as their uncertainty estimates.
Previous approaches have used GPs \cite{deisenroth2011pilco,kamtheDataEfficient2018},
ensembles of neural networks \cite{curiEfficient2020,chuaDeepReinforcementLearning2018}
and variational inference \cite{galImproving2016,houthooftVIME2017}. \todo{Could specify here how the previous methods have used these models}
However, each method has its pros and cons.
In this paper we present a method which combines some of the pros from NNs with the benefits of GPs.

%However, there has been little leveraging
